---
title: "r-python exercise-1 eshaanmathakari"
format: html
editor: visual
Author: Eshaan Mathakari
Description: Performing basic arithmetic and summarising functions in R for Exercise 1 HomeWork
---

# Installing packages
```{r}
#install.packages("pacman")
library(pacman)

p_load(dlookr,
       DMwR2, 
       GGally, # Pair-wise plots using ggplot2
       Hmisc, # Data analysis 
       palmerpenguins, 
       tidyverse)
```

# Loading Datasets into our file
```{r}
data(algae, package = "DMwR2")

algae |> glimpse()
```

## Finding mean
```{r}
algae$a1 |>
  mean()
```

## Finding median
```{r}
algae$a1 |>
  median()
```

## Finding mode
```{r}
Mode <- function(x, na.rm=FALSE){
if(na.rm) x<-x[!is.na(x)]
ux <- unique (x)
return (ux[which.max(tabulate(match(x, ux)))])
}

algae$a2 |> Mode()
```

## Using the centralValue function from DMwR
```{r}
algae$a1 |> centralValue()
algae$speed |> centralValue()
```

## Finding the variance in dataset Algae
```{r}
algae$a1 |> var()
```

## Finding the Standard deviation in dataset Algae
```{r}
algae$a1 |> sd()
```

# Computing basic functions like min, max, range
```{r}
algae$a1 |> range()
algae$a1 |> max()
algae$a1 |> min()
```

## Finding Quartile and InterQuartile range
```{r}
algae$a1 |> IQR()
algae$a1 |> quantile()
algae$a1 |> quantile(probs = c(0.2, 0.8))
```

# Finding the missing values and incomplete rows in Dataset
```{r}
library(dplyr)
library(purrr)
# Compute the total number of NA values in the dataset
nas <- algae %>% 
  purrr::map_dbl(~sum(is.na(.))) %>% 
  sum()

cat("The dataset contains ", nas, "NA values. \n")
incomplete_rows <- algae %>% 
  summarise_all(~!complete.cases(.)) %>%
  nrow()
```

```{r}
cat("The dataset contains ", incomplete_rows, "(out of ", nrow(algae),") incomplete rows. \n")
```

## Summarising the dataset using the summary function
```{r}
algae |> summary()
```

# Using different Describe functions on PalmerPenguins Dataset
```{r}
install.packages("palmerpenguins")
library(palmerpenguins)
data("penguins")
```

## HMISC Describe function
```{r}
penguins |> Hmisc::describe()
```

## DLOOKR Describe function
```{r}
penguins |> dlookr::describe()
```

# Summarising the subsets of data
```{r}
algae |>
  summarise(avgNO3 = mean(NO3, na.rm=TRUE),
            medA1 = median(a1))
```

```{r}
algae |>
  select(mxPH:Cl) |>
  summarise_all(list(mean, median), na.rm = TRUE)
```

```{r}
algae |>
  select(a1:a7) |>
  summarise_all(funs(var))
```

```{r}
algae |>
  select(a1:a7) |>
  summarise_all(c("min", "max"))
```

# Summarising our data by group
```{r}
algae |>
  group_by(season, size) |>
  summarise(nObs = n(), mA7 = median(a7))
penguins |> 
  group_by(species) |>
  summarise(var = var(bill_length_mm, na.rm = TRUE))
```

# Aggregating Data based on their bill length
```{r}
penguins |>
  group_by(species) |>
  reframe(var = quantile(bill_length_mm, na.rm = TRUE))
penguins |>
  group_by(species) |>
  dlookr::describe(bill_length_mm)
```
